---
title: "00_methods paper"
author: "Brenna Kelly"
date: "2025-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r}

library(sf)
library(tmap)
library(ggplot2)
library(leaflet)
library(viridis)
library(ggridges)
library(lubridate)
library(tidyverse)
library(extrafont)
library(tidycensus)
library(colorspace)
library(data.table)
library(RColorBrewer)

loadfonts(quiet = T)

```


## Read in data

```{r}

# o3_files <- paste0("../ebus-python/data/output/o3/", 
#                    list.files("../ebus-python/data/output/o3/"))
# 
# o3_table <- lapply(o3_files, fread)
# 
# o3_df <- do.call("rbind", o3_table)

```

write it / read it:

```{r}

# write.csv(o3_df, "o3_full_period.csv", row.names = FALSE)
o3_df <- fread("o3_full_period.csv")
# pm_df <- fread("output/pm25/all_adj.csv")

```


## pre-preparing

```{r}

# note 9/9: the units were wrong when ran 2019-2024
# - use with transform here for now
# v2 should have the correct units

o3_df_clean <- o3_df |>
  filter(date_time >= "2019-01-01") |>
  # ppm to ppb
  mutate(yhat = yhat * 1000) |>
  # mutate(val = val * 1000) |>
  mutate(adj = val - yhat)

summary(o3_df_clean$date_time)

o3_df_clean$date <- as.Date(o3_df_clean$date_time)
o3_df_clean <- o3_df_clean |>
  filter(val < 175) # see note below

## convert to spatial for mapping
o3_sf <- st_as_sf(o3_df_clean, 
                  coords = c("lon", "lat"), 
                  crs = 4326)


### there are a handful (12) observations with ridiculously high O3
## - these took place at the railyard and (I think?) the depot
## - they happened either around midnight or not in summer
## - will exclude higher than 175
# o3_sf |>
#   # filter(date == "2020-08-14") |>
#   filter(val > 175) |>
#   tm_shape() +
#   tm_dots(fill = "val", size = 2, fill_alpha = 0.9, 
#           tm_scale_continuous(values = "viridis"))


## negative values were truncated at 0
## - is this a reasonable assumption?
aggregate(o3_df_clean$yhat, by = list(o3_df_clean$val == 0), FUN = median)
## probably
# o3_sf |>
#   # filter(date == "2020-08-14") |>
#   filter(val == 0) |>
#   tm_shape() +
#   tm_dots(fill = "yhat", size = 2, fill_alpha = 0.9, 
#           tm_scale_continuous(values = "viridis"))

```

### Add characteristics

```{r}

# day of week
o3_df_clean$day_of_week <- wday(o3_df_clean$date_time)
o3_df_clean$day_of_week_label <- lubridate::wday(o3_df_clean$date_time, label = TRUE)
# weekend
o3_df_clean$day_type <- ifelse(o3_df_clean$day_of_week %in% c(1, 7), 
                           "weekend", "weekday")

# time of day
o3_df_clean$time_of_day <- hour(o3_df_clean$date_time)

# month, season
# create variables of interest
o3_df_clean <- o3_df_clean |>
  mutate(month = lubridate::month(date_time, label = TRUE)) |>
  mutate(day_of_week = lubridate::wday(date_time, week_start = 7)) |> #, # start week with Sunday
                            #label = TRUE)) |>
  mutate(hour = lubridate::hour(date_time)) |>
  mutate(weekend = ifelse(day_of_week %in% c("Sunday", "Saturday"),
                          "weekend", "weekday"))

o3_df_clean <- o3_df_clean |>
  mutate(season = case_when(o3_df_clean$date_time < "2023-03-20" ~ "Winter",
                            o3_df_clean$date_time < "2023-06-21" ~ "Spring",
                            o3_df_clean$date_time < "2023-09-22" ~ "Summer",
                            o3_df_clean$date_time < "2023-12-21" ~ "Autumn",
                            o3_df_clean$date_time >= "2023-12-21" ~ "Winter"))

```


# figures and such  
<!-- - trends over time (2020-2023)   -->
- summary statistics for ebus and monitors  
<!-- - summary of observations; histograms showing data density/sparsity   -->
- histograms  
- map of adjustments over all time (0.025, median, q0.975; mean just for fun)  
- a zoomed-in map with a lot of data; points and aggregate? to show level of detail we get  
- time series for year and a typical (average?) day for each month; to show the diurnal seasonal cycle (and monthly)  

## summary statistics

```{r}

summary(o3_df_clean$adj)

ggplot(o3_df_clean, aes(x = adj)) +
  geom_histogram(bins = 40, fill = "black") +
  labs(x = "Adjustment (ppb)", y = "Count", 
       title = "Distribution of Adjustments",
       subtitle = "Observed value \u2013 predicted value") +
  theme_bw(base_family = "Arial")

summary(o3_df_clean$adj)

# library(ggpubr)
# ggqqplot(o3_df_clean$adj)

# Shapiro-Wilk test for normality
# shapiro.test(o3_df_clean$)

# skewness
library(moments)
skewness(o3_df_clean$adj) # let's call it normal


summary(o3_df_clean$adj)


```

The median adjustment is -1.46 (IQR [-5.44, 1.94]), meaning in a typical scenario in this dataset, the mobile observations overestimated O3 by about 1.5 ppb. Adjustment values also ranged from -80.7 to 144.8 — substantial over- and underestimates, respectively.


```{r}

# potential ones: 
# "2024-07-16"
# "2024-08-08"
### "2020-08-08"
## "2023-04-28"
## "2023-02-08"
### "2024-01-16"

### July 5 > no afternoon, negative adjustments in south I-15
### July 20 > little afternoon, negative adjustments in south I-15

### looking for July, for Daniel
sub_daily_jul <- o3_df_clean |>
  filter(month == "Jul") |>
  mutate(year = lubridate::year(date_time)) |>
  filter(date == "2020-07-20") |>
  filter(year == 2020) |>
  # filter(date == "2023-04-28") |>
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(slc_cnty)) |>
  mutate(morning_afternoon_evening = case_when(time_of_day %in% c(4:11) ~ "1morning",
                                               time_of_day %in% c(12:18) ~ "2afternoon",
                                               time_of_day %in% c(19, 20, 21, 22, 
                                                                  23, 0, 1, 2, 3) ~ "3evening"))

jul_20 <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(sub_daily_jul) +
  tm_dots(fill = "adj", fill_alpha = 0.8, tm_scale_continuous(values = "-brewer.spectral",
                                            midpoint = 0,
                                            limits = c(-53.3, 53.3)),
          fill.legend = tm_legend(show = FALSE)) +
  tm_legend(reverse = TRUE) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_facets(by = c("morning_afternoon_evening"), rows = 1) +
  tm_title("July 20, 2020")


sub_daily_apr <- o3_df_clean |>
  filter(date == "2023-04-28") |>
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(slc_cnty)) |>
  mutate(morning_afternoon_evening = case_when(time_of_day %in% c(4:11) ~ "1morning",
                                               time_of_day %in% c(12:18) ~ "2afternoon",
                                               time_of_day %in% c(19, 20, 21, 22, 
                                                                  23, 0, 1, 2, 3) ~ "3evening"))
# show that it can be different depending on time of day (still visible without this)
sub_daily_apr <- rbind(sub_daily_apr |>
  filter(morning_afternoon_evening == "1morning") |>
  arrange(-adj),
  sub_daily_apr |>
  filter(morning_afternoon_evening != "1morning") |>
  arrange(adj))

apr_28 <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(sub_daily_apr) +
  tm_dots(fill = "adj", fill_alpha = 0.8, tm_scale_continuous(values = "-brewer.spectral",
                                            midpoint = 0,
                                            limits = c(-53.3, 53.3)),
          fill.legend = tm_legend(show = FALSE)) +
  tm_legend(reverse = TRUE) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_facets(by = c("morning_afternoon_evening"), rows = 1) +
  tm_title("Apr 28, 2023")


sub_daily_oct <- o3_df_clean |>
  filter(date == "2022-10-04") |>
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(slc_cnty)) |>
  mutate(morning_afternoon_evening = case_when(time_of_day %in% c(4:11) ~ "1morning",
                                               time_of_day %in% c(12:18) ~ "2afternoon",
                                               time_of_day %in% c(19, 20, 21, 22, 
                                                                  23, 0, 1, 2, 3) ~ "3evening"))
oct_4 <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(sub_daily_oct) +
  tm_dots(fill = "adj", fill_alpha = 0.8, tm_scale_continuous(values = "-brewer.spectral",
                                            midpoint = 0,
                                            limits = c(-53.3, 53.3)),
          fill.legend = tm_legend(show = FALSE)) +
  tm_legend(reverse = TRUE) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_facets(by = c("morning_afternoon_evening"), rows = 1) +
  tm_title("Oct 4, 2022")


sub_daily_aug <- o3_df_clean |>
  filter(date == "2020-08-08") |>
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(slc_cnty)) |>
  mutate(morning_afternoon_evening = case_when(time_of_day %in% c(4:11) ~ "1morning",
                                               time_of_day %in% c(12:18) ~ "2afternoon",
                                               time_of_day %in% c(19, 20, 21, 22, 
                                                                  23, 0, 1, 2, 3) ~ "3evening"))
aug_8 <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(sub_daily_aug) +
  tm_dots(fill = "adj", fill_alpha = 0.8, tm_scale_continuous(values = "-brewer.spectral",
                                            midpoint = 0,
                                            limits = c(-53.3, 53.3)),
          fill.legend = tm_legend(show = FALSE)) +
  tm_legend(reverse = TRUE) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_facets(by = c("morning_afternoon_evening"), rows = 1) +
  tm_title("Aug 8, 2020")


sub_daily_jan <- o3_df_clean |>
  filter(date == "2024-01-16") |>
  st_as_sf(coords = c("lon", "lat"), crs = st_crs(slc_cnty)) |>
  mutate(morning_afternoon_evening = case_when(time_of_day %in% c(4:11) ~ "1morning",
                                               time_of_day %in% c(12:18) ~ "2afternoon",
                                               time_of_day %in% c(19, 20, 21, 22, 
                                                                  23, 0, 1, 2, 3) ~ "3evening"))
jan_16 <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(sub_daily_jan) +
  tm_dots(fill = "adj", fill_alpha = 0.8, tm_scale_continuous(values = "-brewer.spectral",
                                            midpoint = 0,
                                            limits = c(-53.3, 53.3)),
          fill.legend = tm_legend(show = FALSE)) +
  tm_legend(reverse = TRUE) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_facets(by = c("morning_afternoon_evening"), rows = 1) +
  tm_title("Jan 16, 2024")

tmap_arrange(jul_20, oct_4, 
             apr_28, jan_16,#aug_8, 
             ncol = 4)

jul_20

```


###
stop dicking around and make a map

```{r}

o3_sf

## make the grid cells
slc_cnty <- get_acs(geography = "county",
                    variables = c('B01001_001'),
                    state = "UT",
                    geometry = TRUE,
                    year = 2021) |>
  filter(NAME == "Salt Lake County, Utah")

slc_cnty <- st_transform(slc_cnty, st_crs(o3_sf))

fishnet = slc_cnty %>%
  st_make_grid(cellsize = 0.005, what = "polygons") %>% 
  # make cell size 0.005 degrees, or ~500 m
  st_as_sf() %>%
  rename(geometry = x) %>%
  mutate(fishnet_id = 1:n()) %>%
  dplyr::select(fishnet_id, geometry)

## assign points to grid cells
fishnet <- st_transform(fishnet, crs = st_crs(o3_sf))
# only grid cells located in salt lake county
fishnet <- st_join(fishnet, slc_cnty, join = st_intersects) |>
  filter(!is.na(GEOID)) |>
  select(!c(GEOID, NAME, variable, estimate, moe))

grid_sf <- st_join(o3_sf, fishnet, join = st_intersects)

## aggregating data to the grid level)

grid_agg <- grid_sf |>
  st_drop_geometry() |>
  merge(fishnet, by = "fishnet_id") |>
  group_by(fishnet_id) |>
  summarize(obs_mean = mean(val, na.rm = TRUE),
            pred_mean = mean(yhat, na.rm = TRUE),
            adj_mean = mean(adj, na.rm = TRUE),
            obs_median = median(val, na.rm = TRUE),
            pred_median = median(yhat, na.rm = TRUE),
            adj_median = median(adj, na.rm = TRUE),
            obs_low = quantile(val, 0.025, na.rm = TRUE),
            pred_low = quantile(yhat, 0.025, na.rm = TRUE),
            adj_low = quantile(adj, 0.025, na.rm = TRUE),
            obs_high = quantile(val, 0.975, na.rm = TRUE),
            pred_high = quantile(yhat, 0.975, na.rm = TRUE),
            adj_high = quantile(adj, 0.975, na.rm = TRUE),
            count = n()) %>%
  ungroup()

grid_agg_dat <- merge(fishnet, grid_agg, by = "fishnet_id")

tmap_mode(mode = "plot")
### maps

# median
median_adj <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_median", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("Median")

# 97.5th percentile
upper_adj <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_high", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("97.5th percentile")

# 2.5th percentile
lower_adj <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_low", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("2.5th percentile") +
  tmap_options(component.autoscale = TRUE)

tmap_arrange(median_adj, lower_adj, upper_adj, ncol = 3)


### areas with consistently high or low adjustment

grid_agg_dat$adj_consistent <- case_when(grid_agg_dat$adj_low > 0 &
                                           grid_agg_dat$adj_high > 0 ~ "underestimation",
                                         grid_agg_dat$adj_low < 0 &
                                           grid_agg_dat$adj_high < 0 ~ "overestimation")
grid_agg_dat$adj_consistent <- ifelse(is.na(grid_agg_dat$adj_consistent),
                                      "not consistent", grid_agg_dat$adj_consistent)
table(grid_agg_dat$adj_consistent)

tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_median", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = FALSE,
                                      title = "Aggregated \nAdjustment")) +
  tm_shape(grid_agg_dat |>
             filter(adj_consistent != "not consistent")) +
  tm_polygons(fill = "adj_median", col = "black", lwd = 0.5,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = FALSE,
                                      title = "Aggregated \nAdjustment"))

aggregate(grid_agg_dat$count, by = list(grid_agg_dat$adj_consistent), FUN = mean)
# the grid cells with consistent results tend to have fewer than 4 
aggregate(grid_agg_dat$count, by = list(grid_agg_dat$adj_consistent == "not consistent"), FUN = mean)
# in fact, those without a consistent adjustment trend have an average of 7918 level-3 observations
# those with a consistent adjustment trend have an average of 3.5 level-3 observations


```

### Prototypical examples

```{r}

grid_agg_oct <- grid_sf |>
  st_drop_geometry() |>
  # merge(fishnet, by = "fishnet_id") |>
  mutate(month = lubridate::month(date)) |>
  # mutate(day_of_week = lubridate::wday(date, label = TRUE)) |>
  # mutate(hour = lubridate::hour(date)) |>
  filter(month == 3)|># & day_of_week == "Tue" & hour == 7) |>
  group_by(fishnet_id)|>#, month) |>
  summarize(obs_mean = mean(val, na.rm = TRUE),
            pred_mean = mean(yhat, na.rm = TRUE),
            adj_mean = mean(adj, na.rm = TRUE),
            obs_median = median(val, na.rm = TRUE),
            pred_median = median(yhat, na.rm = TRUE),
            adj_median = median(adj, na.rm = TRUE),
            obs_low = quantile(val, 0.025, na.rm = TRUE),
            pred_low = quantile(yhat, 0.025, na.rm = TRUE),
            adj_low = quantile(adj, 0.025, na.rm = TRUE),
            obs_high = quantile(val, 0.975, na.rm = TRUE),
            pred_high = quantile(yhat, 0.975, na.rm = TRUE),
            adj_high = quantile(adj, 0.975, na.rm = TRUE),
            count = n()) %>%
  ungroup()

grid_agg_oct <- merge(fishnet, grid_agg_oct, by = "fishnet_id")
grid_agg_months <- merge(fishnet, grid_agg_months, by = "fishnet_id")

tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_months) +
  tm_polygons(fill = "adj_low", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-53.6, 53.6),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_facets(by = "month")

# tmap_mode(mode = "plot")
### maps
# median
#median_adj <- 
# summary(grid_agg_oct$adj_high)
# summary(grid_agg_jul$adj_high)
# summary(grid_agg_feb$adj_high)
# summary(grid_agg_oct$adj_low)
# summary(grid_agg_jul$adj_low)
# summary(grid_agg_feb$adj_low)

oct <-
  tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_oct) +
  tm_polygons(fill = "adj_low", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-53.6, 53.6),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("October")

jul <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_jul) +
  tm_polygons(fill = "adj_low", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-53.6, 53.6),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("July")
feb <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_feb) +
  tm_polygons(fill = "adj_low", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-53.6, 53.6),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("February")

tmap_arrange(feb, jul, oct, ncol = 3)


oct <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_oct) +
  tm_polygons(fill = "count", lwd = 0,
              tm_scale_continuous_log10(values = "viridis",
                                  limits = c(1, 59300)),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("October")
jul <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_jul) +
  tm_polygons(fill = "count", lwd = 0,
              tm_scale_continuous_log10(values = "viridis",
                                  limits = c(1, 59300)),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("July")
feb <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_feb) +
  tm_polygons(fill = "count", lwd = 0,
              tm_scale_continuous_log10(values = "viridis",
                                  limits = c(1, 59300)),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("February")

tmap_arrange(feb, jul, oct, ncol = 3)




# 97.5th percentile
upper_adj <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_high", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-53.6, 53.6),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("97.5th percentile")

# 2.5th percentile
lower_adj <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_low", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      title = "Aggregated \nAdjustment")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_title("2.5th percentile") +
  tmap_options(component.autoscale = TRUE)

tmap_arrange(median_adj, lower_adj, upper_adj, ncol = 3)


### areas with consistently high or low adjustment

grid_agg_dat$adj_consistent <- case_when(grid_agg_dat$adj_low > 0 &
                                           grid_agg_dat$adj_high > 0 ~ "underestimation",
                                         grid_agg_dat$adj_low < 0 &
                                           grid_agg_dat$adj_high < 0 ~ "overestimation")
grid_agg_dat$adj_consistent <- ifelse(is.na(grid_agg_dat$adj_consistent),
                                      "not consistent", grid_agg_dat$adj_consistent)
table(grid_agg_dat$adj_consistent)

tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(grid_agg_dat) +
  tm_polygons(fill = "adj_median", lwd = 0,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = FALSE,
                                      title = "Aggregated \nAdjustment")) +
  tm_shape(grid_agg_dat |>
             filter(adj_consistent != "not consistent")) +
  tm_polygons(fill = "adj_median", col = "black", lwd = 0.5,
              tm_scale_continuous(values = c("royalblue1", 
                                             "white", 
                                             "orangered"),
                                  limits = c(-43, 43),
                                  midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = FALSE,
                                      title = "Aggregated \nAdjustment"))

aggregate(grid_agg_dat$count, by = list(grid_agg_dat$adj_consistent), FUN = mean)
# the grid cells with consistent results tend to have fewer than 4 
aggregate(grid_agg_dat$count, by = list(grid_agg_dat$adj_consistent == "not consistent"), FUN = mean)
# in fact, those without a consistent adjustment trend have an average of 7918 level-3 observations
# those with a consistent adjustment trend have an average of 3.5 level-3 observations


```




### number of raw observations / L3 obs & raw observations / grid cell

```{r}

mob <- fread("data/mobile/mobile_o3_v5.csv")

o3_mob <- merge(o3_df_clean, mob[, c("times", "LON",
                               "LAT", "MOB")], 
                by.x = c("date_time", "lon", "lat"),
                by.y = c("times", "LON", "LAT"))

# join to grid_shape, for join to fishnet
head(o3_mob)
head(grid_sf)

o3_mob_shp <- merge(o3_mob, 
                    st_drop_geometry(grid_sf[, c("date_time", "normalized_east",
                                                 "normalized_north", "fishnet_id")]),
                    by = c("date_time", "normalized_east", "normalized_north"))

o3_mob_shp <- o3_mob_shp |>
  merge(fishnet, by = "fishnet_id") |>
  group_by(fishnet_id) |>
  summarize(count = sum(MOB)) |>
  # summarize(count = n()) |> # don't use this; get the actual full count of obs
  ungroup()

o3_mob_shp <- merge(fishnet, o3_mob_shp, by = "fishnet_id")

mob_mon_map <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(st_as_sf(o3_mob_shp)) +
  tm_polygons(fill = "count", lwd = 0,
              tm_scale_intervals(values = "viridis",
                                 n = 10, style = "quantile"),
                                  # limits = c(-43, 43),
                                  # midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = FALSE, # change for single figure / panel
                                      title = "Count of mobile\nobservations \n(deciles)")) +
  tm_layout(frame = FALSE, legend.frame = FALSE)


# make a panel with stationary monitor locations
aqs_sites <- fread("data/aqs/aqs_o3_v5.csv") |>
  distinct(site.num, .keep_all = TRUE) |>
  st_as_sf(coords = c("longitude", "latitude"),
           crs = 4326)

fread("data/aqs/aqs_o3_v5.csv") |>
  filter(day_time >= "2019-01-01") |>
  distinct(day_time) |>
  nrow()
# 52588

stat_mon_map <- tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(aqs_sites) +
  tm_dots(size = 1, shape = 18, fill = "black",
          fill.legend = tm_legend(reverse = TRUE)) +
  tm_text(text = "site.num", size = 0.8, col = "black", xmod = 0.9, ymod = 0.1,
          options = opt_tm_text(point.label.gap = 0,
                                just = "left")) +
  tm_layout(frame = FALSE, legend.frame = FALSE)

tmap_arrange(stat_mon_map, mob_mon_map)

# plotted together
tm_shape(slc_cnty) +
  tm_borders(fill_alpha = 0) +
  tm_shape(st_as_sf(o3_mob_shp)) +
  tm_polygons(fill = "count", lwd = 0,
              tm_scale_intervals(values = "viridis",
                                 n = 10, style = "quantile"),
              # limits = c(-43, 43),
              # midpoint = 0),
              fill.legend = tm_legend(reverse = TRUE, show = TRUE,
                                      # change for single figure / panel
                                      title = "Count of mobile\nobservations \n(deciles)")) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_shape(aqs_sites) +
  tm_dots(size = 0.75, shape = 18, fill = "red",
          fill.legend = tm_legend(reverse = TRUE))


```





https://cran.r-project.org/web/packages/tableone/vignettes/introduction.html

```{r}
 



# summary(
#   kruskal.test(adj ~ season, data = o3_df_clean)
# )
# 
# aov_test <- kruskal.test(adj ~ season, data = o3_df_clean)
# aov_test
# 
# TukeyHSD(aov(adj ~ as.factor(time_of_day), 
#              data = o3_df_clean))
# 
# tukey.test <- TukeyHSD(aov(adj ~ as.factor(time_of_day), 
#                            data = o3_df_clean))
# tukey.test


# CreateTableOne(vars = "adj", strata = c("month", "day_of_week", "hour", "weekend",
#                                         "season"), data = o3_df_clean,
#                testNonNormal = kruskal.test)


aggregate(o3_df_clean$adj, by = list(o3_df_clean$month),
          FUN = 'quantile', probs = c(0.025, 0.25, 0.5, 0.75, 0.975))

```

### Plot and summarize adjustments by time of day
### summarize adjustments across season

```{r}


# aggregate(o3_df_clean$val, by = list(o3_df_clean$))

agg_time <- cbind(
  aggregate(o3_df_clean$val, by = list(o3_df_clean$time_of_day), 
            FUN = quantile, probs = c(0.05)) |>
    dplyr::rename(time = Group.1,
           low_O3 = x),
  aggregate(o3_df_clean$val, by = list(o3_df_clean$time_of_day), 
            FUN = quantile, probs = c(0.5)) |>
    select(x) |>
    dplyr::rename(median_O3 = x),
  aggregate(o3_df_clean$val, by = list(o3_df_clean$time_of_day), 
            FUN = quantile, probs = c(0.95)) |>
    select(x) |>
    dplyr::rename(high_O3 = x),
  aggregate(o3_df_clean$yhat, by = list(o3_df_clean$time_of_day), 
            FUN = quantile, probs = c(0.05)) |>
    select(x) |>
    dplyr::rename(low_O3_pred = x),
  aggregate(o3_df_clean$yhat, by = list(o3_df_clean$time_of_day), 
            FUN = quantile, probs = c(0.5)) |>
    select(x) |>
    dplyr::rename(median_O3_pred = x),
  aggregate(o3_df_clean$yhat, by = list(o3_df_clean$time_of_day), 
            FUN = quantile, probs = c(0.95)) |>
    select(x) |>
    dplyr::rename(high_O3_pred = x))

agg_time_long <- agg_time |>
  pivot_longer(cols = c("low_O3", "median_O3", "high_O3",
                        "low_O3_pred", "median_O3_pred", "high_O3_pred"),
               names_to = "variable")


ggplot(agg_time_long, aes(x = time, y = value, group = variable)) +
  geom_line(aes(colour = variable), lwd = 0.75, alpha = 0.9) +
  scale_color_manual(values = c("orangered4", "orangered",
                                "navy", "royalblue",
                                #"maroon3", "violet",
                                "gray20", "gray60")) +
  scale_x_continuous(expand = c(0, 0), 
                     breaks = seq(from = 0, to = 23, by = 2)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 70)) +
  xlab("Hour of Day") +
  ylab("Ozone concentration (ppb)") +
  theme_bw()



## hourly boxplots of observed and predicted
o3_df_clean |>
  pivot_longer(cols = c("val", "yhat"), 
               names_to = "o3_measure") |>
  mutate(hour = as.factor(hour)) |>
  ggplot(aes(x = hour, y = value, col = factor(o3_measure))) +
  geom_boxplot(alpha = 0.2) +
  # scale_colour_manual(values = c("slateblue4",
  #                                "red")) +
  geom_boxplot(outlier.shape = NA) +
  scale_colour_manual(values = c("lightseagreen",
                                 "red")) +
  scale_y_continuous(expand = c(0, 0))+#, limits = c(0, 175)) +
  scale_x_discrete(expand = c(0, 0), 
                   breaks = seq(from = 0, to = 23, by = 2)) +
  xlab("Hour of Day") +
  ylab("Ozone concentration (ppb)") +
  theme_bw()

## and the adjustments
overall_boxplot <- o3_df_clean |>
  mutate(hour = lubridate::hour(date_time),
         hour = as.factor(hour)) |>
  ggplot(aes(x = hour, y = adj)) +#, col = as.factor(month))) +
  geom_boxplot(alpha = 0.1) +
  geom_boxplot(outlier.shape = NA) +
  scale_colour_manual(values = "black") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), 
                   breaks = seq(from = 0, to = 23, by = 2)) +
  xlab("Hour of Day") +
  ylab("Adjustment (ppb)") +
  ggtitle("Adjustment Distribution") +
  theme_bw()# +
  # facet_wrap(~ month)

median_for_boxplot <- o3_df_clean |>
  mutate(hour = lubridate::hour(date_time),
         hour = as.factor(hour)) |>
  group_by(hour) |>
  summarize(median = median(adj)) |>
  ggplot(aes(x = hour, y = median)) +
  geom_point() +
  # scale_y_continuous(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0),
                     breaks = seq(from = 0, to = 23, by = 2)) +
  xlab("Hour of Day") +
  ylab("Adjustment (ppb)") +
  ggtitle("Median Adjustment") +
  theme_bw()

library(patchwork)

median_for_boxplot / overall_boxplot +
   plot_layout(heights = c(2, 5))



```



```{r}

```




```{r}

# exceptional events
# https://aqs.epa.gov/aqsweb/documents/codetables/qualifiers.html
aqs <- read.csv("data/aqs/o3/hourly_44201_2023.csv")
2.1
table(aqs$Qualifier)
# Just FYI, MDL = 0.005, or 5; minimum detectable level

aqs <- aqs |>
  mutate(qualifier_def = as.character(
    case_when(Qualifier == "5" ~ "outlier",
              Qualifier == "IF" ~ "fire, canada",
              Qualifier == "IT" ~ "fire, us"))) |>
  mutate(date_time = ymd_hm(paste(Date.Local, Time.Local))) |>
  mutate(date_time = force_tz(date_time, tzone = "MST")) |>
  mutate(date = ymd(Date.Local)) |>
  mutate(date = force_tz(date, tzone = "MST"))

exceptions <- as.data.frame(
  table(aqs$date, aqs$qualifier_def)) |>
  filter(Freq > 0) |>
  rename(date = Var1,
         exception = Var2) |>
  select(!Freq)

o3_df_clean$date <- as.factor(date(o3_df_clean$date_time))

# merge to add exceptions to o3 data
o3_df_clean <- merge(o3_df_clean, exceptions, by = "date", all.x = TRUE)
table(o3_df_clean$exception)

o3_df_clean$exception <- ifelse(is.na(o3_df_clean$exception),
                                "no exception", as.character(o3_df_clean$exception))

# adjustment
ggplot(o3_df_clean, aes(x = adj)) +
  geom_histogram(bins = 40, aes(fill = exception)) +
  labs(x = "Adjustment (ppb)", y = "Count", 
       title = "Distribution of Adjustments",
       subtitle = "Observed value \u2013 predicted value") +
  theme_bw(base_family = "Century Gothic") +
  facet_wrap(~ exception, scales = "free_y")

# observed values
ggplot(o3_df_clean, aes(x = val)) +
  geom_histogram(bins = 40, aes(fill = exception)) +
  labs(x = "O3 concetration (ppb)", y = "Count", 
       title = "Distribution of Observed O3",
       subtitle = "Mobile observations") +
  theme_bw(base_family = "Century Gothic") +
  facet_wrap(~ exception, scales = "free_y")


aggregate(o3_df_clean$adj, by = list(o3_df_clean$exception), FUN = mean)
aggregate(o3_df_clean$adj, by = list(o3_df_clean$exception), FUN = sd)
# on average, we overestimate O3 more during fires but less during "outliers"
# but this is highly variable

# anova
res_aov <- aov(adj ~ exception, data = o3_df_clean)

hist(res_aov$residuals) # looks normal

summary(res_aov) # adjustments are different across exception

# interestingly, O3 tends to be higher on days without exceptional events
aggregate(o3_df_clean$val, by = list(o3_df_clean$exception), FUN = median)
aggregate(o3_df_clean$val, by = list(o3_df_clean$exception), FUN = mean)
# weird that "outlier" days have such low O3

## just checking that we captured the right dates; apparently we did
# we, meaning me and my computer
outlier <- o3_df_clean |>
  filter(exception == "outlier")

unique(date(outlier$date_time))

outlier_aqs <- aqs |>
  filter(qualifier_def == "outlier")

unique(date(outlier_aqs$date_time))

?CreateTableOne




summary(o3_df_clean)


```


## maps of adjustments  

```{r}

```


## zoomed-in map with a lot of data, to show resolution

```{r}

```




## daily time series for seasons

```{r}

```




# extras  
- how adjustments change residential exposure  
    - using landscan global 2023  
    - using census block groups  
    
```{r}



```


```{r}

# get census geometry, population data

# v20 <- load_variables(2020, "dhc", cache = TRUE)

# View(v20)

slc_cnty <- get_decennial(geography = "block group",
                          variables = c('H6_001N'),
                          state = "UT",
                          geometry = TRUE,
                          year = 2020,
                          sumfile = "dhc") |>
  mutate(county = str_sub(GEOID, start = 0, end = 5)) |>
  filter(county ==  "49035")

# join adjustments to census block groups

tm_shape(slc_cnty) +
  tm_polygons(fill = "gray90", col = "white") +
  tm_shape(o3_sf) +
  tm_dots(fill_alpha = 0.05, fill = "orangered")

```


```{r}
# join and aggregate

o3_sf <- st_transform(o3_sf, crs = st_crs(slc_cnty))

o3_slc <- st_join(o3_sf, slc_cnty, st_intersects)

# aggregate to block group (0.025, median, 0.975), then join to block group geometry

block_counts <- as.data.frame(table(o3_slc$GEOID))

o3_slc_agg <- aggregate(o3_slc$adj, by = list(o3_slc$GEOID), FUN = quantile,
                       probs = c(0.025, 0.5, 0.975)) |>
  mutate(lo_adj = x[, 1]) |>
  mutate(med_adj = x[, 2]) |>
  mutate(hi_adj = x[, 3]) |>
  rename(GEOID = `Group.1`) |>
  select(!c(x))

o3_slc_agg <- merge(slc_cnty, o3_slc_agg, by = "GEOID", all.x = TRUE)

```

## time series for full year

```{r}

# o3_slc

```


## daily time series for month

```{r}

library(tidybayes)

m_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
names(m_labels) <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)

o3_slc |>
  st_drop_geometry() |>
  group_by(month(date_time), hour(date_time)) |>
  median_qi(val, yhat, .width = 0.95) |>
  rename(hour = `hour(date_time)`,
         month = `month(date_time)`) |>
  # filter(month == 10) |>
  ggplot(aes(x = hour, y = val, ymin = val.lower, ymax = val.upper)) +
  geom_line(colour = "red") +
  geom_ribbon(aes(fill = "Mobile"), alpha = 0.5) +
  geom_line(aes(x = hour, y = yhat), colour = "navy") +
  geom_ribbon(aes(ymin = yhat.lower, ymax = yhat.upper,
                  fill = "Stationary"),
              alpha = 0.35) +
  scale_fill_manual(labels = c("Mobile", "Stationary"), values = c("hotpink", "purple")) +
  labs(x = "Hour of Day", y = "Adjustment (ppm)",
       fill = "Data") +
  facet_wrap(~ month, labeller = labeller(month = m_labels)) +
  theme_minimal()

```


```{r}
# map
o3_slc_agg_empty <- o3_slc_agg |>
  filter(is.na(med_adj))

# pop centroids for large areas
pop <- read.csv("methods paper/data/CenPop2020_Mean_BG49.csv") |>
  mutate(GEOID = paste0(STATEFP, str_pad(COUNTYFP, pad = "0", 
                                         width = 3, side = "left"), 
                        str_pad(TRACTCE, pad = "0", 
                                width = 6, side = "left"), BLKGRPCE))
o3_slc_agg$area <- as.numeric(
  st_area(o3_slc_agg))
large <- o3_slc_agg |>
  filter(!is.na(med_adj)) |>
  filter(area >= 50000000)
large_pop <- pop[which(pop$GEOID %in% large$GEOID), ] |>
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

# map of median adjustments for 2023, aggregated to block group
tm_shape(o3_slc_agg) +
  tm_polygons(fill = "med_adj", style = "cont", 
              palette = "-spectral", lwd = 0, title = "Median \nAdjustment") +
  tm_shape(o3_slc_agg_empty) +
  tm_polygons(fill = "gray90", lwd = 0) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_shape(large_pop) +
  tm_dots(fill = "gray20") +
  tm_add_legend(labels = "Population \nCenters",
                type = "symbols", fill = "gray20",
                size = 0.25)

# low adjustments
tm_shape(o3_slc_agg) +
  tm_polygons(fill = "lo_adj", style = "cont", 
              palette = "-brewer.spectral", lwd = 0, title = "Q0.025 \nAdjustment") +
  tm_shape(o3_slc_agg_empty) +
  tm_polygons(fill = "gray90", lwd = 0) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_shape(large_pop) +
  tm_dots(fill = "gray20") +
  tm_add_legend(labels = "Population \nCenters",
                type = "symbols", fill = "gray20",
                size = 0.25)

# high adjustments
tm_shape(o3_slc_agg) +
  tm_polygons(fill = "hi_adj", style = "cont",# fill_alpha = 0.25, 
              palette = "-brewer.spectral", lwd = 0, title = "Q97.5 \nAdjustment") +
  tm_shape(o3_slc_agg_empty) +
  tm_polygons(fill = "gray90", lwd = 0) +
  tm_layout(frame = FALSE, legend.frame = FALSE) +
  tm_shape(large_pop) +
  tm_dots(fill = "gray20") +
  tm_add_legend(labels = "Population \nCenters",
                type = "symbols", fill = "gray20",
                size = 0.25)

tmap_mode(mode = "plot")

```

preliminary investigation

```{r}

tm_shape(o3_slc_agg) +
  tm_polygons(fill = "value", style = "cont", palette = "viridis", lwd = 0) +
  tm_layout(frame = FALSE, legend.frame = FALSE)

cor.test(o3_slc_agg$value, o3_slc_agg$med_adj)
# weak, no correlation

# at a finer resolution; within block groups
cor.test(o3_slc$adj, o3_slc$value)
# slight negative correlation between population and adjustment
# i.e., the air quality tends to be better than we think in areas with higher population

# let's test population density
o3_slc_agg$pop_density <- o3_slc_agg$value / (o3_slc_agg$area / 1000) # people / sq km

cor.test(o3_slc_agg$pop_density, o3_slc_agg$med_adj)
# slight negative association, approaches significance

# at a finer resolution with population density
pop_adj_test <- merge(o3_slc, st_drop_geometry(o3_slc_agg), by = "GEOID")

cor.test(pop_adj_test$pop_density, pop_adj_test$adj)
# very slight positive correlation; significant but high N

tm_shape(o3_slc_agg) +
  tm_polygons(fill = "pop_density", style = "cont", palette = "viridis", lwd = 0) +
  tm_layout(frame = FALSE, legend.frame = FALSE)


# sampling observations, weighted by population density

o3_slc_agg_comp <- o3_slc_agg |>
  filter(!is.na(lo_adj))

set.seed(4178)

nsample = 100
samp_idx <- sample(seq_len(nrow(o3_slc_agg_comp)), nsample, 
                   prob = o3_slc_agg_comp$pop_density)
o3_slc_sample <- o3_slc_agg_comp[samp_idx, ]


```



hypothetical residential exposure profiles

```{r}

# select a random set of block groups, weighted by number of observations (not pop density)

n_obs <- o3_slc |>
  group_by(GEOID) |>
  st_drop_geometry() |>
  count()

```
```{r}
# weighting by n obs in a block group
samp_idx <- sample(seq_len(nrow(n_obs)), 100, 
                   prob = n_obs$n)
```
```{r}
# getting sampled block groups
sampled_geoids <- n_obs[samp_idx, "GEOID"]
# selecting observations in sampled block groups
o3_slc_sample <- o3_slc[which(o3_slc$GEOID %in% sampled_geoids$GEOID), ]

```
```{r}
# select just one; to walk through
case_study <- o3_slc_sample |>
  filter(GEOID == 490351025012)

# time series
case_study_long <- rbind(case_study[, c("date_time", "val", "GEOID")] |>
                           rename(measure = val) |>
                           mutate(stat = "val"),
                         case_study[, c("date_time", "yhat", "GEOID")] |>
                           rename(measure = yhat) |>
                           mutate(stat = "yhat"))
```
```{r}
ggplot(case_study_long, aes(x = date_time, 
                            y = measure, group = stat,
                            colour = stat)) +
  geom_line()
```
```{r}
# time series by month
val_by_month <- aggregate(case_study$val, by = list(hour(case_study$date_time),
                                    month(case_study$date_time)), FUN = median) |>
  # as.data.frame() |>
  filter(Group.2 == 10) |>
  ggplot(aes(x = Group.1, y = x)) +
  geom_line()
```
```{r}
yhat_by_month <- aggregate(case_study$yhat, by = list(hour(case_study$date_time),
                                    month(case_study$date_time)), FUN = median) |>
  # as.data.frame() |>
  filter(Group.2 == 10) |>
  ggplot(aes(x = Group.1, y = x)) +
  geom_line()
```
```{r}
  #
# val_by_month <- val_by_month |>
#   filter(Group.2 == 10)
# yhat_by_month <- yhat_by_month |>
#   filter(Group.2 == 10)
# 
# ggplot(val_by_month, aes(x = Group.1, y = x)) +
#   geom_line() +
#   geom_line(data = yhat_by_month, colour = "red")
```
```{r}
library(ggdist)
library(tidybayes)

m_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
names(m_labels) <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)

case_study |>
  st_drop_geometry() |>
  group_by(month(date_time), hour(date_time)) |>
  median_qi(val, yhat, .width = 0.95) |>
  rename(hour = `hour(date_time)`,
         month = `month(date_time)`) |>
  # filter(month == 10) |>
  ggplot(aes(x = hour, y = val, ymin = val.lower, ymax = val.upper)) +
  geom_lineribbon(fill = "hotpink", colour = "red", alpha = 0.5) +
  geom_lineribbon(aes(x = hour, y = yhat, 
                      ymin = yhat.lower, ymax = yhat.upper),
                  fill = "purple", colour = "navy", alpha = 0.5) +
  labs(x = "Adjustment (ppm)", y = "Hour of Day") +
  facet_wrap(~ month, labeller = labeller(month = m_labels)) +
  theme_minimal()
```
```{r}

# geom_line(aes(y=y, x=x, colour = "sin"))+
#       geom_ribbon(aes(ymin=lower, ymax=upper, x=x, fill = "band"), alpha = 0.3)+
  

case_study |>
  st_drop_geometry() |>
  group_by(month(date_time), hour(date_time)) |>
  median_qi(val, yhat, .width = 0.95) |>
  rename(hour = `hour(date_time)`,
         month = `month(date_time)`) |>
  # filter(month == 10) |>
  ggplot(aes(x = hour, y = val, ymin = val.lower, ymax = val.upper)) +
  geom_line(colour = "red") +
  geom_ribbon(aes(fill = "Mobile"), alpha = 0.5) +
  geom_line(aes(x = hour, y = yhat), colour = "navy") +
  geom_ribbon(aes(ymin = yhat.lower, ymax = yhat.upper,
                  fill = "Stationary"),
              alpha = 0.35) +
  scale_fill_manual(labels = c("Mobile", "Stationary"), values = c("hotpink", "purple")) +
  labs(x = "Hour of Day", y = "Adjustment (ppm)",
       fill = "Data") +
  facet_wrap(~ month, labeller = labeller(month = m_labels)) +
  theme_minimal()

# ?median_qi
# 
# case_study_agg |>
#   filter(Group.2 == 10) |>
#   ggplot(aes(x = Group.1, y = x)) +
#   stat_summary(fun.data = )
# 
# 
# ggplot(case_study, aes(x = date_time, y = val)) +
#   stat_summary(geom = "ribbon",
#                fun.data = mean_cl_normal,
#                fun.args = list(conf.int = 0.95))
#   geom_line(stat = "summary", fun.y = "median") +
#   geom_ribbon(stat = "summary", ymin = function(x) quantile(x, 0.25), ymax = function(x) quantile(x, 0.75), fill = "lightblue", alpha = 0.5)


```


## adjustments over time

```{r}

o3_df$date_time <- ymd_hms(o3_df$date_time)

o3_df$year <- as.factor(
  year(o3_df$date_time))

###### aside: remove last two weeks of 2019
o3_df <- o3_df[which(o3_df$year != 2019), ]
######

o3_df$quarter <- as.factor(
  quarter(o3_df$date_time))

o3_df$month <- as.factor(
  month(o3_df$date_time))

table(o3_df$month,
      o3_df$year)

o3_df$hour <- as.factor(
  hour(o3_df$date_time))

o3_df$day_of_week <- as.factor(
  wday(o3_df$date_time))

##### aside: convert to ppm
# o3_df$adj <- o3_df$adj
# don't

# distribution of adjustments over hour
ggplot(o3_df, aes(x = hour, y = adj, fill = hour)) + 
  geom_boxplot() +
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "Year, Season", y = "Adjustment")

table(o3_df$adj > 0.2,
      o3_df$hour) # most of the wildly high adjustments are late in the night

test <- o3_df |>
  filter(adj < 0.06 &
           adj > -0.06)

q_labels <- c("Jan-Mar", "Apr-Jun",
              "Jul-Sep", "Oct-Dec")
names(q_labels) <- c(1, 2, 3, 4)

ggplot(test, aes(x = adj, y = hour, fill = stat(x))) + 
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.001, show.legend = NA) +
  # coord_cartesian(clip = "on") +
  scale_fill_viridis_c(option = "D") +
  coord_flip() +
  labs(x = "Adjustment (ppm)", y = "Hour of Day") +
  theme_minimal() +
  facet_wrap(~ quarter, labeller = labeller(quarter = q_labels))


# does variability in adjustments change by month-time of day
m_labels <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
names(m_labels) <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12)

ggplot(test, aes(x = adj, y = hour, fill = stat(x))) + 
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.001, show.legend = NA) +
  # coord_cartesian(clip = "on") +
  scale_fill_viridis_c(option = "D") +
  coord_flip() +
  labs(x = "Adjustment (ppm)", y = "Hour of Day") +
  theme_minimal() +
  facet_wrap(~ month, labeller = labeller(month = m_labels))

test_2020 <- test |>
  filter(year == 2020)

ggplot(test_2020, aes(x = adj, y = hour, fill = stat(x))) + 
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.001, show.legend = NA) +
  # coord_cartesian(clip = "on") +
  scale_fill_viridis_c(option = "D") +
  coord_flip() +
  labs(x = "Adjustment (ppm)", y = "Hour of Day") +
  theme_minimal() +
  facet_wrap(~ month, labeller = labeller(month = m_labels))



# is October's variability explained by week day-hour
w_labels <- c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat")
names(w_labels ) <- c(1, 2, 3, 4, 5, 6, 7)

test_oct <- test |>
  filter(year == 2020 &
           month == 10)

ggplot(test_oct, aes(x = adj, y = hour, fill = stat(x))) + 
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.001, show.legend = NA) +
  # coord_cartesian(clip = "on") +
  scale_fill_viridis_c(option = "D") +
  coord_flip() +
  labs(x = "Adjustment (ppm)", y = "Hour of Day") +
  theme_minimal() +
  facet_wrap(~ day_of_week, labeller = labeller(day_of_week = w_labels))

# hourly for entire month
test_oct$day_hour <- as.factor(
  round_date(test_oct$date_time, unit = "hour"))

ggplot(test_oct, aes(x = adj, y = day_hour, fill = stat(x))) + 
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.001, show.legend = NA, lwd = 0.25) +
  # coord_cartesian(clip = "on") +
  scale_fill_viridis_c(option = "D") +
  coord_flip() +
  labs(x = "Adjustment (ppm)", y = "Hour of Day") +
  theme_minimal()# +
  #facet_wrap(~ week)#, labeller = labeller(day_of_week = w_labels))

# did october have a unique spatial extent?
test_no_oct <- test |>
  filter(year == 2020) |>
  filter(month != 10)

max(test_no_oct$lon) - min(test_no_oct$lon)
max(test_no_oct$lat) - min(test_no_oct$lat)
max(test_oct$lon) - min(test_oct$lon)
max(test_oct$lat) - min(test_oct$lat)

# distribution of adjustments over year, season
ggplot(o3_df, aes(x = quarter, y = adj, fill = year)) + 
  geom_boxplot() +
  scale_fill_brewer(palette = "Dark2") +
  labs(x = "Year, Season", y = "Adjustment")


# have adj changed over time?
adj_over_time <- group_by(o3_df, year, month) %>%
  summarise(
    mean = mean(adj, na.rm = TRUE),
    sd = sd(adj, na.rm = TRUE)
  )

plot(adj_over_time$mean)

# do adjustments change over time of day?
# have adj changed over time?
adj_over_tod <- group_by(o3_df, hour) %>%
  summarise(
    mean = mean(adj, na.rm = TRUE),
    sd = sd(adj, na.rm = TRUE)
  )

plot(adj_over_tod$mean)

# with ridges, for more information
ggplot(o3_df, aes(x = adj, y = month,#paste(year, month))
                    fill = stat(x))) + 
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.001) +
  coord_cartesian(clip = "on") +
  scale_fill_viridis_c(option = "C")
  # geom_density_ridges()
# change axis limits? might show more color variability

ggplot(o3_df, aes(x = adj, y = year, fill = stat(x))) +
  geom_density_ridges_gradient() +
  scale_fill_viridis_c(name = "Depth", option = "C") 


# probably can't interpret anova when years are not independent
oneway.test(adj ~ year,
  data = o3_df,
  var.equal = FALSE
)

# has o3 changed over time?
o3_over_time <- group_by(o3_df, year, month) %>%
  summarise(
    mean = mean(yhat, na.rm = TRUE),
    sd = sd(yhat, na.rm = TRUE)
  )
# increase in average and SD

plot(o3_over_time$mean)


```

```{r}

slc_cnty #<- st_transform(slc_cnty, st_crs(out_sf))

fishnet = slc_cnty %>%
  st_make_grid(cellsize = 0.005, what = "polygons") %>% # make cell size 0.005 degrees, or ~500 m
  st_as_sf() %>%
  rename(geometry = x) %>%
  mutate(fishnet_id = 1:n()) %>%
  dplyr::select(fishnet_id, geometry)

## assign points to grid cells
# dat_sf <- st_as_sf(dat, coords = c("longitude", "latitude"), crs = 4326)

fishnet <- st_transform(fishnet, crs = st_crs(o3_sf))

grid_sf <- st_join(o3_sf, fishnet, join = st_intersects)

```

## summary of observations

Count

```{r}
library(lubridate)
# by hour of day
grid_sf$time_mt <- with_tz(grid_sf$date_time, "America/Denver")

grid_sf$time_of_day <- hour(grid_sf$time_mt)

ggplot(grid_sf, aes(x = time_of_day)) +
  geom_histogram(bins = 24)
```
```{r}
# by year
grid_sf$year <- year(grid_sf$time_mt)

ggplot(grid_sf, aes(x = year)) +
  geom_histogram(stat = "count") +
  labs(x = "Year")

```
```{r}
# by month
grid_sf$month <- lubridate::month(grid_sf$time_mt)
grid_sf$month_label <- lubridate::month(grid_sf$time_mt, label = TRUE)
```
```{r}
ggplot(grid_sf, aes(x = month_label)) +
  geom_histogram(stat = "count") +
  labs(x = "Month")

# by day of week
grid_sf$day_of_week <- wday(grid_sf$time_mt)
grid_sf$day_of_week_label <- lubridate::wday(grid_sf$time_mt, label = TRUE)

grid_sf$day_type <- ifelse(grid_sf$day_of_week %in% c(1, 7), 
                           "weekend", "weekday")
```
```{r}
ggplot(grid_sf, aes(x = day_of_week_label)) +
  geom_histogram(stat = "count") +
  labs(x = "Day of Week")

# by hour of day
# grid_sf$time_mt <- with_tz(grid_sf$time_mt, "America/Denver")
# 
# grid_sf$time_of_day <- hour(grid_sf$time_mt)
# 
# ggplot(grid_sf, aes(x = time_of_day)) +
#   geom_histogram(bins = 24)

```


```{r}

# 0.070 ppm exceedance and adjustments

# is the predicted o3 above 0.070, and is this an underestimate?
exc_adj = data.frame(table(o3_df$yhat > 0.07, o3_df$adj > 0))

below_under = exc_adj[1, "Freq"]
above_under = exc_adj[2, "Freq"]
below_over = exc_adj[3, "Freq"]
above_over = exc_adj[4, "Freq"]

prop.table(table(o3_df$yhat > 0.07, o3_df$adj > 0))
# - 37.1% of the time, O3 is thought to be below 0.070 AND this is an underestimate
##    worse than safe
# - 64.4% of the time, O3 is thought to be below 0.070 AND this is an overestimate
##    better than safe
# - 0.1% of the time, O3 is thought to be above 0.070 AND this is an underestimate
##    worse than unsafe
# - 0.4% of the time, O3 is thought to be above 0.070 AND this is an overestimate
##    better than unsafe

# i.e., in relative terms
below_under / above_under
# a <0.070 underestimate is 167-times more likely than a >0.070 underestimate
below_over / above_over
# a <0.070 overestimate is 377-times more likely than a >0.070 overestimate
(below_under + above_under) / (below_over + above_over)
# overall, an overestimate is 69% more likely than an underestimate
(below_under + below_over) / (above_under + above_over)
# overall, a <0.070 prediction is 211-times more likely than >0.070
table(o3_df$val > 0.07)[1] / table(o3_df$val > 0.07)[2]
# a <0.070 mobile observation is 259-times more likely than a >0.070 observation

## note: this does not tell us about the spread, or variability; i.e., how wrong we are; just the direction

## range of over- / under-estimate
summary(o3_df[which(o3_df$adj < 0), "val"])
# hist(o3_df[which(o3_df$adj < 0), "val"])
# when o3 is better than expected, it's typically "off" by 24 ppb
summary(o3_df[which(o3_df$adj > 0), "val"])
# hist(o3_df[which(o3_df$adj > 0), "val"])
# when o3 is worse than expected, it's typically "off" by 35 ppb
o3_under <- o3_df[which(o3_df$adj > 0), ] # getting all underestimates
summary(o3_under$yhat)
# when o3 is worse than expected, the conc is usually 29 ppb, but as high as 100 ppb
o3_under[which(o3_under$yhat > 0.099), ]
# - this was the afternoon of July 12, 2021, throughout the valley; monitors thought it was under 100 ppb, but it was as high as 109 ppb
bad_locs <- o3_under[which(o3_under$yhat > 0.099), ] |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

tmap_mode("plot")

tm_shape(bad_locs) +
  tm_dots(col = "val", palette = "-Spectral", style = "cont",
          size = 0.1, alpha = 0.6, border.lwd = 0)

bad_day <- o3_df[which(date(o3_df$date_time) == "2021-07-12" &
                          hour(o3_df$date_time) > 12 &
                          hour(o3_df$date_time) < 17), ] |>
  st_as_sf(coords = c("lon", "lat"), crs = 4326) |>
  arrange(desc(date_time))

summary(bad_day$val)

bad_day_val <- tm_shape(bad_day) +
  tm_symbols(col = "val", shape = 19, #fill.scale = tm_scale_continuous(limits = c(0.035, 0.114)),
             palette = "-brewer.spectral", style = "cont", breaks = seq(0.06, 0.11, by = 0.015),
             size = 0.75, alpha = 0.8,
             # title = "O3 (mobile)",
          tm_add_legend("col", type = "symbols",
                 title = "title here" 
                 )) +
  tm_layout(title = "O3 (mobile)", title.position = c('center', 'top'), frame = FALSE)
  # tm_borders(lwd = 0)

bad_day_yhat <- tm_shape(bad_day) +
  tm_symbols(col = "yhat", shape = 19, #fill.scale = tm_scale_continuous(limits = c(0.035, 0.114)),
             palette = "-brewer.spectral", style = "cont", breaks = seq(0.06, 0.11, by = 0.015),
             size = 0.75, alpha = 0.8,
             # title = "O3 (mobile)",
          tm_add_legend("col", type = "symbols",
                 title = "title here" 
                 )) +
  tm_layout(title = "O3 (stationary)", title.position = c('center', 'top'), frame = FALSE)
  # tm_borders(lwd = 0)
#
tmap_arrange(bad_day_val, bad_day_yhat)


summary(o3_df[which(o3_df$adj < 0), "val"])
# hist(o3_df |>
#        filter(o3_df$adj < 0) |>
#        select("val") |>
#        mutate(val = as.numeric(val)))

summary(o3_df[which(o3_df$adj < 0), "yhat"])
# hist(o3_df[which(o3_df$adj < 0), "yhat"])

summary(o3_df$yhat)
summary(o3_df$val)


o3_df_long <- o3_df |>
  pivot_longer(cols = c("val", "yhat")) |>
  mutate(name = ifelse(name == "val", "Mobile", "Stationary")) |>
  arrange(name)

ggplot(o3_df_long, aes(x = value, colour = name, fill = name)) +
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.005) +
  scale_color_manual(values = c("#E7298AFF", "#D95F02FF")) +
  scale_fill_manual("O3 Data", values = c("#E7298AFF", "#D95F02FF")) +
  labs(x = "O3 (ppm)", y = "Count") +
  # labs(fill = 'NEW LEGEND TITLE') +
  theme_minimal()


ggplot(o3_df, aes(x = val, fill = "#E7298AFF")) +
  geom_histogram(position = "identity", alpha = 1, binwidth = 0.01) +
  # scale_color_manual(values = c("#E7298AFF")) +
  # scale_fill_manual("O3 Data", values = c("#E7298AFF", "#D95F02FF")) +
  labs(x = "O3 (ppm)", y = "Count") +
  # labs(fill = 'NEW LEGEND TITLE') +
  theme_minimal()

```



Get SL County geometry, grid cells

```{r}

slc_cnty <- st_read("/Users/brenna/Downloads/tl_2022_us_county/tl_2022_us_county.shp") |>
  filter(STATEFP == "49" & COUNTYFP == "035")

fishnet = slc_cnty %>%
  st_make_grid(cellsize = 0.01, what = "polygons") %>%
  st_as_sf() %>%
  rename(geometry = x) %>%
  mutate(fishnet_id = 1:n()) %>%
  dplyr::select(fishnet_id, geometry)

```


```{r}

# preparing data for spatial join
dat_sf <- st_as_sf(o3_df, coords = c("lon", "lat"), crs = 4326)
dat_sf$longitude <- dat_sf$lon
dat_sf$latitude <- dat_sf$lat

fishnet <- st_transform(fishnet, crs = st_crs(dat_sf))

grid_sf <- st_join(dat_sf, fishnet, join = st_intersects)

# aggregating data to the grid level
grid_no_geom <- grid_sf |>
  st_drop_geometry() |>
  merge(fishnet, by = "fishnet_id") |>
  group_by(fishnet_id) |>
  summarize(obs_mean = mean(val, na.rm = TRUE),
            pred_mean = mean(yhat, na.rm = TRUE),
            adj_mean = mean(adj, na.rm = TRUE),
            obs_median = median(val, na.rm = TRUE),
            pred_median = median(yhat, na.rm = TRUE),
            adj_median = median(adj, na.rm = TRUE),
            obs_high = quantile(val, 0.95, na.rm = TRUE),
            pred_high = quantile(yhat, 0.95, na.rm = TRUE),
            adj_high = quantile(adj, 0.95, na.rm = TRUE),
            count = n()) %>%
  ungroup()

grid_dat <- merge(fishnet, grid_no_geom, by = "fishnet_id")

```

Mapping the aggregation to grid cell level

```{r}

# peak adjustment
summary(grid_dat$adj_high)

pal <- colorNumeric(
  palette = "Spectral",
  domain = c(-0.024, 0.024), reverse = TRUE)

leaflet() |>
  setView(-111.86, 40.675, zoom = 10) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(data = grid_dat, color = ~pal(adj_high), fillOpacity = 0.75, stroke = FALSE) |>
  addLegend('bottomright', pal = pal, values = grid_dat$adj_high,
            title = "Peak adjustment")

# observed median
summary(grid_dat$obs_median)

pal <- colorNumeric(
  palette = "viridis",
  domain = c(0, 0.0662), reverse = FALSE)

leaflet() |>
  setView(-111.86, 40.675, zoom = 10) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(data = grid_dat, color = ~pal(obs_median), fillOpacity = 0.75, stroke = FALSE) |>
  addLegend('bottomright', pal = pal, values = grid_dat$obs_median,
            title = "Median observed value")

# predicted median
summary(grid_dat$pred_median)

pal <- colorNumeric(
  palette = "viridis",
  domain = c(0, 0.0662), reverse = FALSE)

leaflet() |>
  setView(-111.86, 40.675, zoom = 10) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(data = grid_dat, color = ~pal(pred_median), fillOpacity = 0.75, stroke = FALSE) |>
  addLegend('bottomright', pal = pal, values = grid_dat$pred_median,
            title = "Median predicted value")

```





```{r}

cor.test(grid_sf$val, grid_sf$yhat)
# correlated 88% of the time

grid_corr <- grid_sf |>
  st_drop_geometry() |>
  group_by(hour) |>
  summarise(correlation = cor(val, yhat),
            count = n())

summary(grid_corr$correlation)

cor.test(grid_corr$correlation, grid_corr$count)

plot(grid_corr$hour, grid_corr$count)

grid_sf |>
  st_drop_geometry() |>
  group_by(hour) |>
  summarise(correlation = cor(val, yhat)) |>
  plot()


cor.test(grid_sf$val, grid_sf$yhat)

grid_cor <- grid_sf |>
  st_drop_geometry() |>
  merge(fishnet, by = "fishnet_id") |>
  group_by(fishnet_id) |>
  summarize(obs_mean = mean(val, na.rm = TRUE),
            pred_mean = mean(yhat, na.rm = TRUE),
            adj_mean = mean(adj, na.rm = TRUE),
            obs_median = median(val, na.rm = TRUE),
            pred_median = median(yhat, na.rm = TRUE),
            adj_median = median(adj, na.rm = TRUE),
            obs_high = quantile(val, 0.95, na.rm = TRUE),
            pred_high = quantile(yhat, 0.95, na.rm = TRUE),
            adj_high = quantile(adj, 0.95, na.rm = TRUE),
            correlation = cor(val, yhat),
            count = n()) %>%
  ungroup() |>
  as.data.frame()

pal <- colorNumeric(
  palette = "RdBu",
  domain = c(-1, 1), reverse = FALSE)

test <- grid_cor |>
  filter(!is.na(correlation))

leaflet() |>
 setView(-111.86, 40.675, zoom = 10) |>
 addProviderTiles(providers$CartoDB.Positron) |>
 addPolygons(data = grid_dat, color = ~pal(grid_cor$correlation), fillOpacity = 0.9, stroke = FALSE) |>
 addLegend('bottomright', pal = pal, values = grid_cor$correlation,
           title = "Correlation between \n observed and predicted O3", na.label = "Too few observations")

leaflet() |>
  setView(-111.86, 40.675, zoom = 10) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(data = grid_dat, color = ~pal(correlation), fillOpacity = 0.75, stroke = FALSE) |>
  addLegend('bottomright', pal = pal, values = grid_cor$correlation,
            title = "Median predicted value", na.label = "Too few observations")

```





Aggregating to grid cell-year-month

```{r}

grid_no_geom_ym <- grid_sf |>
  st_drop_geometry() |>
  merge(fishnet, by = "fishnet_id") |>
  group_by(fishnet_id, year, month) |>
  summarize(obs_mean = mean(val, na.rm = TRUE),
            pred_mean = mean(yhat, na.rm = TRUE),
            adj_mean = mean(adj, na.rm = TRUE),
            obs_median = median(val, na.rm = TRUE),
            pred_median = median(yhat, na.rm = TRUE),
            adj_median = median(adj, na.rm = TRUE),
            obs_high = quantile(val, 0.95, na.rm = TRUE),
            pred_high = quantile(yhat, 0.95, na.rm = TRUE),
            adj_high = quantile(adj, 0.95, na.rm = TRUE),
            count = n()) %>%
  ungroup()

grid_dat_ym <- merge(fishnet, grid_no_geom_ym, by = "fishnet_id")

# for October 2020

# observed median

oct_20 <- grid_dat_ym |>
  filter(#year == 2020 &
           month == 10)

# aggregate(grid_dat_ym$obs_median, by = list(grid_dat_ym$year, grid_dat_ym$month), FUN = median)
summary(oct_20$obs_median)

pal <- colorNumeric(
  palette = "viridis",
  domain = c(0, 0.05609), reverse = FALSE)

leaflet() |>
  setView(-111.86, 40.675, zoom = 10) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(data = oct_20, color = ~pal(obs_median), fillOpacity = 0.75, stroke = FALSE) |>
  addLegend('bottomright', pal = pal, values = oct_20$obs_median,
            title = "Median observed value")

# predicted median
summary(oct_20$pred_median)

pal <- colorNumeric(
  palette = "viridis",
  domain = c(0, 0.05609), reverse = FALSE)

leaflet() |>
  setView(-111.86, 40.675, zoom = 10) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(data = oct_20, color = ~pal(pred_median), fillOpacity = 0.75, stroke = FALSE) |>
  addLegend('bottomright', pal = pal, values = oct_20$pred_median,
            title = "Median observed value")

```

## looped animation for monthly adjustments

```{r}

summary(grid_dat_ym$adj_high)

library(mapview)
library(webshot)
install.packages("webshot")


for(i in unique(grid_dat_ym$year)) {
  for(j in unique(grid_dat_ym$month)) {
    
    dat_loop <- grid_dat_ym |>
      filter(year == i &
               month == j)
    
    pal <- colorNumeric(
      palette = "Spectral",
      domain = c(-0.05, 0.05), reverse = TRUE)
    
    adj_map <- leaflet() |>
      setView(-111.86, 40.675, zoom = 10) |>
      addProviderTiles(providers$CartoDB.Positron) |>
      addPolygons(data = dat_loop, color = ~pal(adj_high), fillOpacity = 0.75, stroke = FALSE) |>
      addLegend('bottomright', pal = pal, values = dat_loop$adj_high,
                title = "Median adjustment")
    
    # Define filename
    filename <- paste0("adj_map_", i, "_", j, ".png")
    
    # Save map to image
    mapshot(adj_map, file = filename)
    
  }
}


```




# things we don't need the full dataset for  
- archetypes (year, month, DOW, hour)


Display a typical day (y = o3, x = time) for a given grid cell on a weekday in a month in a year

```{r}

ggplot(agg_time_long, aes(x = time, y = value, group = variable)) +
  geom_line(aes(colour = variable)) +
  scale_color_manual(values = c("#E7298A", "#e3909c",
                                "#1B9E77", "#a0be95", 
                                "#D95F02", "#ecb56a"))

```





